Title

### 2024-06-27 paper/main.tex
Updates

\begin{enumerate}
	\item \textbf{Trustworthiness: } Although most algorithms have
		asymptotic theory proving them to be consistent, their finite
		sample properties are not well understood. In real-world
		settings with finite samples, these algorithms can make obvious
		mistakes, for example, the constraint-based algorithms are
		biased towards sparse graphs \citep{Malinsky2024}.
		Consequently, researchers often have rely on post-hoc
		modifications based on model testing methods and their domain
		knowledge.
	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple causal
		structures can be faithful to a given observational dataset.
		Thus, using only observational data, these algorithms can only
		recover the MECs that can contain unoriented edges. However,
		most methods for downstream tasks, such as identification or
		causal effect estimation, assume knowledge of a fully oriented
		DAG. As a result, researchers have to rely on their expertise
		to orient these edges or make further assumptions to make use
		of pairwise orientation techniques.
	\item \textbf{Algorithm and hyperparameter selection:} Each algorithm
		has the same goal but they can learn very different models
		depending on the choice of algorithm and their hyperparameters.
		For example, the choice of the conditional independence (CI)
		test being used for constraint-based method depends on the
		dataset, similarly for the scoring method for score-based
		methods. With no obvious way to test the impact/performance of
		these choices on a given dataset, it becomes quite hard to
		chose the optimal algorithm for a given dataset.
\end{enumerate}

As a result of these challenges, doesn't matter if the researcher is manually
constructing the DAG or using an algorithm, they need to do some manual
modifications to the model before being able to use them for analysis. One of
the common ways to guide this modification is to use Conditional Independence
(CI) based model testing \cite{Ankan2021}. As DAGs imply CI statements that can
be tested in the given data, and then model can be modified based on the
results of this test. Figure~\ref{fig:ci_testing} shows an example of testing
using CIs. This approach has two main drawbacks:
\begin{enumerate}
	\item CIs are implied by missing edges only, and hence we can not test
		whether an incorrect edge is present in the model.
	\item For larger graphs, the number of CI tests can be very large, for
		example, the alarm network with $ 37 $ nodes has $ 287 $
		implied CIs. It can be difficult to look at all these tests for
		violations.
\end{enumerate}

@@

On the other hand, in Structural Equation Models (SEMs) literature, more
emphasis has been placed on helping researchers manually construct models while
integrating their expertise. Researchers typically develop an initial model
based on domain knowledge and use various tools to guide modifications to
improve the fit of their model to data while it still making theoretical sense.
This model construction process is known as \emph{Specification Search}. A
common method for specification search is \emph{modification indices}, which
identify potential changes to the model structure that improve the model fit
and ranks them using the most improvement to the model fit. This ranking allows
the researchers to focus first on the suggestions that can improve the model
the most and can use their expertise to choose the most optimal modifications.

@@

% One of the drawbacks of these global fit metrics is that they do not only test
% for the structure but also the fitted parameterization.

@@

				\draw[->]  (age) to[bend right=20] (ed);
				\draw[->]  (sex) to (ed);
				\draw[->]  (age) to (hrpw);
				\draw[->]  (ed) to (hrpw);
				\draw[->]  (sex) to (hrpw);
				\draw[->]  (ed) to (income);
				\draw[->]  (hrpw) to (income);
				\draw[->]  (occ) to[out=300, in=0, looseness=1.4] (income.east);
				\draw[->]  (sex) to (income);
				\draw[->]  (ed) to[out=90, in=150, looseness=1.3] (occ);
				\draw[->]  (sex) to (occ);	
			\end{scope}
			\begin{scope}[xshift=-3cm]
				\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
				\node[rectangle, align=center, inner sep=1pt] at (5, 0) {
					\begin{minipage}{.2\textwidth}
						\begin{equation*}
							\begin{split}
								\textnormal{Age} &\ci \textnormal{Incm} \rvert \textnormal{Edct, Hrpw, Sex} \\
								\textnormal{Age} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
								\textnormal{Age} &\ci \textnormal{Sex} \\
								\textnormal{Hrpw} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
							\end{split}
						\end{equation*}
					\end{minipage}
					};
			\end{scope}
		\end{tikzpicture}
		\caption{\todo[inline]{Show a full example of testing using this model}}
		\label{fig:ci_testing}
\end{figure}

The rest of the paper is structured as follows. In
Section~\ref{sec:background}, we give a background on the measures of
association for various data types. In Section~\ref{sec:mixed_association}, we
present our method for computing mixed data association. In
Section~\ref{sec:modification}, we present how to use this method to rank the
modifications, and lastly in Section~\ref{sec:empirical}, we show some
empirical analysis on how well this modification process works.

@@

\section{Measure of Association for Mixed Data}
\label{sec:mixed_association}

@@

\section{Using measure of association for DAG construction}
\label{sec:modification}

@@

Using this mixed data partial association measure, we propose an iterative DAG
modification / construction procedure. Users can start with an empty graph.
Given a DAG $ G $ and a dataset $ D $, for each pair for variable $ X $ and $ Y
$ in $ G $ that does not have an edge between them, we compute the partial
association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. The measure of
assocation in this case can be interpreted as the unexplained observed
association between them given the current model. We can then use this to rank
pair of variables which have very high unexplained association between them.
Based on the domain expertise, the user can then choose one of these suggested
pair of variables, decide the direction of the edge between the variables, and
add it to the model. Once the new edge is added we can recompute the assocation
of all other variables and $ X $ and $ Y $.

Additionally, in each iteration of modification, for each pair of variable in $
G $, that has an edge between them in we compute the partial association
measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $. This
measure of assocation can be interpreted as the strength of the edge between
the variables. If the value of this assocation below a certain threshold, we
can remove the edge.

\begin{theorem}
	If we have access to an oracle that can give correct edge directions, taking a
	greedy approach with the produre above results in the correct graph
	being recovered.
\end{theorem}

\section{Web Tool}
\label{sec:web}
Based on the procedure outlined in the previous section, we built a web-based
tool that allows users to interactively create/modify their model. The user
starts by uploading their dataset and optionally the initial DAG. If no initial
DAG is provided an empty graph is initialized. The tool also computes the
Shipley's C value at any point. The web-tool would automatically show all
unexplained assocations in the DAG by adding red edges between them. The edge
weight of the edge represents how high this strength is. Users can also set the
threshold for p-value and association to control how many edges are shown at a
time. Based on these shown potential edges, the user can select the edge that
they want to add and decide the direction for it. If the tool detects an
incorrect edge in the graph, it is shown in black. Once the user is happy with
the graph they have they can export it into different formats to use it for
further computation.

>>>>

\begin{enumerate}
	\item \textbf{Trustworthiness: } Although most algorithms have
		asymptotic theory proving them to be consistent, their finite
		sample properties are not well understood. Moreover, depending
		on the selected hyperparameters of the algorithms, the output
		can vary very significantly. In real-world settings with finite
		samples, these algorithms can make obvious mistakes, for
		example, the constraint-based algorithms are biased towards
		sparse graphs \citep{Malinsky2024}. Consequently, researchers
		often have rely on post-hoc modifications based on model
		testing methods and their domain knowledge.
	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple causal
		structures can be faithful to a given observational dataset.
		Thus, using only observational data, these algorithms can only
		recover the MECs that can contain unoriented edges. However,
		most methods for downstream tasks, such as identification or
		causal effect estimation, assume knowledge of a fully oriented
		DAG. As a result, researchers have to rely on their expertise
		to orient these edges or make further assumptions to make use
		of pairwise orientation techniques.
	% \item \textbf{Algorithm and hyperparameter selection:} Each algorithm
	% 	has the same goal but they can learn very different models
	% 	depending on the choice of algorithm and their hyperparameters.
	% 	For example, the choice of the conditional independence (CI)
	% 	test being used for constraint-based method depends on the
	% 	dataset, similarly for the scoring method for score-based
	% 	methods. With no obvious way to test the impact/performance of
	% 	these choices on a given dataset, it becomes quite hard to
	% 	chose the optimal algorithm for a given dataset.
\end{enumerate}

As a result of these challenges, doesn't matter if the researcher is manually
constructing the DAG or using an algorithm, they need to do some manual
modifications to the model before being able to use them for analysis. One of
the common ways to guide this modification is to use Conditional Independence
(CI) based model testing \cite{Ankan2021}. As DAGs imply CI statements that can
be tested in the given data, and then model can be modified based on the
p-values of these test. Figure~\ref{fig:ci_testing} shows an example of testing
using CIs. This approach has two main drawbacks:
\begin{enumerate}
	\item CIs are implied by missing edges only, and hence we can not test
		whether an incorrect edge is present in the model.
	\item For larger graphs, the number of CI tests can be very large, for
		example, the alarm network with $ 37 $ nodes has $ 287 $
		implied CIs. It can be difficult to look at all these tests for
		violations.
\end{enumerate}

@@

On the other hand, in Structural Equation Models (SEMs) literature, more
emphasis has been placed on helping researchers manually construct models while
integrating their expertise. Researchers typically develop an initial model
based on domain knowledge and use various tools to guide modifications to
improve the fit of their model to data while it still making theoretical sense.
This model construction process is known as \emph{Specification Search}. A
common method for specification search is \emph{modification indices}, which
identify potential changes to the model structure that improve the model fit
and ranks them using the most improvement to the model fit. This ranking allows
the researchers to focus first on the suggestions that can improve the model
the most and while integrating their expertise to choose the most optimal
modifications.

@@

@@

			\draw[->]  (age) to[bend right=20] (ed);
			\draw[->]  (sex) to (ed);
			\draw[->]  (age) to (hrpw);
			\draw[->]  (ed) to (hrpw);
			\draw[->]  (sex) to (hrpw);
			\draw[->]  (ed) to (income);
			\draw[->]  (hrpw) to (income);
			\draw[->]  (occ) to[out=300, in=0, looseness=1.4] (income.east);
			\draw[->]  (sex) to (income);
			\draw[->]  (ed) to[out=90, in=150, looseness=1.3] (occ);
			\draw[->]  (sex) to (occ);	
		\end{scope}
		\begin{scope}[xshift=-3cm]
			\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
			\node[rectangle, align=center, inner sep=1pt] at (5, 0) {
				\begin{minipage}{.2\textwidth}
					\begin{equation*}
						\begin{split}
							\textnormal{Age} &\ci \textnormal{Incm} \rvert \textnormal{Edct, Hrpw, Sex} \\
							\textnormal{Age} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
							\textnormal{Age} &\ci \textnormal{Sex} \\
							\textnormal{Hrpw} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
						\end{split}
					\end{equation*}
				\end{minipage}
				};
		\end{scope}
		\end{tikzpicture}
		\caption{\todo[inline]{Finish figure to show an example of model testing}}
		\label{fig:ci_testing}
\end{figure}
\begin{figure}
	\caption{\todo[inline]{Placeholder for figure showing an example of modification indices}}
\end{figure}

The rest of the paper is structured as follows. In
Section~\ref{sec:background}, we give a background on the measures of
association for various data types. In Section~\ref{sec:mixed_association}, we
present our method for computing mixed data association. In
Section~\ref{sec:modification} and Section~\ref{sec:}, we present how to use
this method to make modifications and the web tool, and lastly in
Section~\ref{sec:empirical}, we show some empirical analysis on how well this
modification process works.

@@

\section{A Measure of Association for Mixed Data}
\label{sec:mixed_association}

@@

\section{Using Measure of Association for DAG Modification and Construction}
\label{sec:modification}

@@

Using this mixed data partial association measure, we propose an iterative DAG
modification procedure. 
\todo[inline]{How similar is this to GES algorithm}
If the user desires to construct the model manually, they can also start with
an empty graph. Given a DAG $ G $ and a dataset $ D $, for each pair for
variable $ X $ and $ Y $ in $ G $ that does not have an edge between them, we
compute the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $.
The measure of association in this case can be interpreted as the unexplained
observed association between them given the current model. We can then use this
to rank pair of variables which have very high unexplained association between
them. Based on the domain expertise, the user can then choose one of these
suggested pair of variables, decide the direction of the edge between the
variables, and add it to the model. Once the new edge is added we can recompute
the association of all other variables and $ X $ and $ Y $.

Additionally, in each iteration of modification, for each pair of variable in $
G $, that has an edge between them in we compute the partial association
measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $. This
measure of association can be interpreted as the strength of the edge between
the variables. If the value of this association below a certain threshold, we
can remove the edge.

\begin{theorem}
	If we have access to an oracle that can give correct edge directions, taking a
	greedy approach with the procedure above results in the correct graph
	being recovered.
\end{theorem}

\section{Web Tool}
\label{sec:web}
Based on the procedure outlined in the previous section, we built a web-based
tool that allows users to interactively create/modify their model. The user
starts by uploading their dataset and optionally the initial DAG. If no initial
DAG is provided an empty graph is initialized. The tool also computes the
Shipley's C value at any point. The web-tool would automatically show all
unexplained associations in the DAG by adding red edges between them. The edge
weight of the edge represents how high this strength is. Users can also set the
threshold for p-value and association to control how many edges are shown at a
time. Based on these shown potential edges, the user can select the edge that
they want to add and decide the direction for it. If the tool detects an
incorrect edge in the graph, it is shown in black. Once the user is happy with
the graph they have they can export it into different formats to use it for
further computation.
